---
title: "Prediction Assignment"
author: "Juhaninho"
date: "25 Juli 2019"
output: html_document
---

## Overview

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har

The goal of this project is to predict the manner in which they did the exercise.

```{r message=FALSE}
library(caret)
library(rpart)
library(randomForest)
```

## Dataset

Loading the training set and test set.

``` {r}
train <- read.csv("data/pml-training.csv", header = TRUE, na.strings=c("NA",""))
test <- read.csv("data/pml-testing.csv", header = TRUE, na.strings=c("NA",""))
```

Removing missing values.

``` {r}
train <- train[, colSums(is.na(train)) == 0]
```

``` {r}
str(train)
```

Removing variables that are irrelevant for the model

``` {r}
irrelevant_variables <- c(
  "X", "user_name", "raw_timestamp_part_1", "raw_timestamp_part_2",
  "cvtd_timestamp", "new_window", "num_window"
)
train <- train[ , -which(names(train) %in% irrelevant_variables)]
```

## Model

Splitting the training data into a 70% training set and a 30% validation set

```{r}
set.seed(8)
trainset <- createDataPartition(train$classe, p=0.7, list = FALSE)
train <- train[trainset,]
valid <- train[-trainset,]
```

Training the model with a decision tree and a random forest classification. Also a 5-fold cross-validation is used to validate the random forest model.

```{r}
dTreeModel <- rpart(classe ~ ., data=train, method="class")
```

```{r}
cv <- trainControl(method="cv", number=5, verboseIter=FALSE)
rfModel <- randomForest(classe ~ ., data = train, ntree = 5, trControl = cv, proximity = TRUE)
```

Validate the model by predicting the values for the validation set

```{r}
cmValidDT <- confusionMatrix(predict(dTreeModel, valid, type = "class"), valid$classe)
cmValidDT$table
```

```{r}
cmValidRF <- confusionMatrix(predict(rfModel, valid), valid$classe)
cmValidRF$table
```

## Prediction

Choosing the random forest classification because of its higher accuracy.
Making predictions for the test set

```{r}
predictions <- predict(rfModel, test)
print(predictions)
```
